Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-29 sports_analytics.Rmd
added to expected points

Researchers have also applied the notion of expected points to evaluate other aspects of the game of football, including quarterback performance and coaching decision.

@White2002 present a tiered logistic regression method that can be, in general, applied to any regression setting with a polychotomous response.
Using this technique, they estimate the value of NFL plays using a simple expected points model with down, yards to go, and yards to goal as predictors.
Accordingly, the model results can be used to obtain ratings and rankings for NFL passers.

@@

The BTM can be implemented in `R` [@R-base] via the \pkg{BradleyTerry2} package [@R-BradleyTerry2]. 
As an example, we consider the data given in @agresti2018introduction (page 247) on tennis results from 2014--2018 for five men's professional players: Novak Djokovic, Roger Federer, Andy Murray, Rafael Nadal, and Stan Wawrinka. 
We fit a BTM to estimate the win probability for each pair of players and obtain a ranking for this group of five.

@@

Perhaps motivated by his success in the Kaggle March Madness competition, Michael Lopez joined the National Football League and launched the [Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/) in 2019.
This annual competition has similarly fueled new research directions in American football and a [special issue on player tracking data in the National Football League (NFL)](https://www.degruyter.com/journal/key/jqas/16/2/html) of *JQAS* [@lopez2020bigger].
Successful entries and their corresponding publications [@DeshpandeEvans2020; @Yurko2020going; @ChuReyersThomsonWu2020] have launched the careers of several of the most prominent early-career researchers in sports analytics.

>>>>

Researchers have also applied the notion of expected points to investigate other aspects of the game of football, including quarterback performance and coaching decision.

As for quarterback evaluation, @White2002 present a tiered logistic regression method that can be, in general, applied to any regression setting with a polychotomous response.
Using this technique, they estimate the value of NFL plays using a simple expected points model with down, yards to go, and yards to goal as predictors.
Accordingly, the model results can be used to obtain ratings and rankings for NFL passers.

Expected point value is also the main component of a novel quarterback evaluation metric introduced by @Reyers2021quarterback. 
The authors take advantage of player tracking data to account for different passing and running options available to the quarterback.
The expected points and probabilities associated with the available options are estimated using the method of ensemble learning via stacking.

@@

The BTM can be implemented in R via the \pkg{BradleyTerry2} package [@R-BradleyTerry2]. 
As an example, we consider the data given in @agresti2018introduction (page 247) on tennis results from 2014--2018 for five men's professional players: Novak Djokovic, Roger Federer, Andy Murray, Rafael Nadal, and Stan Wawrinka. 
We fit a BTM to estimate the win probability for each pair of players and obtain a ranking for this group of five.

@@

Perhaps motivated by his success in the Kaggle March Madness competition, Michael Lopez joined the National Football League and launched the [Big Data Bowl](https://operations.nfl.com/gameday/analytics/big-data-bowl/) in 2019.
This annual competition has similarly fueled new research directions in American football and a [*JQAS* special issue on player tracking data in the National Football League (NFL)](https://www.degruyter.com/journal/key/jqas/16/2/html) of [@lopez2020bigger].
Successful entries and their corresponding publications [@DeshpandeEvans2020; @Yurko2020going; @ChuReyersThomsonWu2020] have launched the careers of several of the most prominent early-career researchers in sports analytics.
